---
---title: "PA1_Template"
---author: "xchassa"
---date: "Sunday, May 17, 2015"
---output: html_document
---

# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r loaddata}
unzip(zipfile="activity.zip")
activityData <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
totalSteps <- tapply(activityData$steps, activityData$date, FUN=sum)
qplot(totalSteps, binwidth=1000, xlab="total number of steps taken each day")
reportMean <- mean(totalSteps, na.rm=TRUE)
reportMedian <- median(totalSteps, na.rm=TRUE)
```

The `mean` is `r reportMean` and the `median` is `r reportMedian`.

## What is the average daily activity pattern?
```{r}
averages <- aggregate(x=list(steps=activityData$steps), by=list(interval=activityData$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")

maxInterval <- averages[which.max(averages$steps),1]
```

The 5-minute interval, on average across all the days in the data set, containing the maximum number of steps is `r maxInterval`.

## Imputing missing values

Total number of missing values calculation :

```{r how_many_missing}
missingVals <- sum(is.na(activityData$steps))
```

The total number of `missing values` is `r missingVals`.

```{r}
naFill <- function(data, pervalue) {
        na_index <- which(is.na(data$steps))
        na_replace <- unlist(lapply(na_index, FUN=function(idx){
                interval = data[idx,]$interval
                pervalue[pervalue$interval == interval,]$steps
        }))
        fill_steps <- data$steps
        fill_steps[na_index] <- na_replace
        fill_steps
}

rdataFill <- data.frame(  
        steps = naFill(activityData, averages),  
        date = activityData$date,  
        interval = activityData$interval)
str(rdataFill)
```

Check that we have filled the NA values :

```{r}
sum(is.na(rdataFill$steps))
```
## Histogram of the total number of steps taken each day

```{r}
fillStepsPerDay <- aggregate(steps ~ date, rdataFill, sum)
colnames(fillStepsPerDay) <- c("date","steps")

##plotting the histogram
ggplot(fillStepsPerDay, aes(x = steps)) + 
       geom_histogram(fill = "purple", binwidth = 1000) + 
        labs(title="Histogram of Steps Taken per Day", 
             x = "Number of Steps per Day", y = "Number of times in a day(Count)") + theme_bw() 
```

Mean and median total number of steps taken per day :
```{r}
stepsMeanFill   <- mean(fillStepsPerDay$steps, na.rm=TRUE)
stepsMedianFill <- median(fillStepsPerDay$steps, na.rm=TRUE)
```

* The imputed data mean is `r stepsMeanFill`
* The imputed data median is `r stepsMedianFill`

Calculate difference between imputed and non-imputed data.
```{r}
meanDiff <- stepsMeanFill - reportMean
medDiff <- stepsMedianFill - reportMedian
```
* The difference between the non-imputed mean and imputed mean is `r meanDiff`
* The difference between the non-imputed mean and imputed mean is `r medDiff`

## Are there differences in activity patterns between weekdays and weekends?

```{r}
Sys.setlocale("LC_TIME", "English")
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
rdataFill$dow = as.factor(ifelse(is.element(weekdays(as.Date(rdataFill$date)),weekdays), "Weekday", "Weekend"))

stepsByInterval <- aggregate(steps ~ interval + dow, rdataFill, mean)

library(lattice)

xyplot(stepsByInterval$steps ~ stepsByInterval$interval|stepsByInterval$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```
